---
title: "DATA 621 - Homework 4"
author: "Vanita Thompson, David Moste, Sadia Perveen"
date: "April 13, 2021"
output:
  pdf_document: default
  word_document: default
  tidy: yes
  html_document:
    highlight: pygments
    theme: cerulean
code_folding: hide
---

In this homework assignment, you will explore, analyze and model a data set containing approximately 8000 records representing a customer at an auto insurance company. Each record has two response variables. The first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.

Your objective is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car. You can only use the variables given to you (or variables that you derive from the variables provided). Below is a short description of the variables of interest in the data set:


```{r setup, include=FALSE}
install.packages("tidyverse", "psych", )
library(tidyverse)
library(psych)
library(corrplot)
library(RColorBrewer)
library(knitr)
library(MASS)
library(caret)
library(kableExtra)
library(ResourceSelection)
library(pROC)
library(kableExtra)
library(reshape2)
library(Amelia)
library(corrplot)

install.packages("mice")
library(mice)

insurance <- read.csv("https://raw.githubusercontent.com/sperveen/DATA621/main/insurance-training.csv", header = TRUE, sep = ",")

insurance_evaluation <- read.csv("https://raw.githubusercontent.com/sperveen/DATA621/main/insurance-evaluation.csv", header = TRUE, sep = ",")


```



## 1. Data Exploration:


```{R}
names(insurance)
str(insurance)
dim(insurance)
kable(summary(insurance))

# The data needs to be cleaned up. We have some variables with $ and some variables with Z_ that needs to be removed. 

insurance$MSTATUS <- gsub('z_', '', insurance$MSTATUS)
insurance$SEX <- gsub('z_', '', insurance$SEX)
insurance$EDUCATION <- gsub('z_', '', insurance$EDUCATION)
insurance$JOB <- gsub('z_', '', insurance$JOB)
insurance$CAR_TYPE <- gsub('z_', '', insurance$CAR_TYPE)
insurance$URBANICITY <- gsub('z_', '', insurance$URBANICITY)
insurance$INCOME <- gsub('[\\$,]', '', insurance$INCOME)
insurance$HOME_VAL <- gsub('[\\$,]', '', insurance$HOME_VAL)
insurance$BLUEBOOK <- gsub('[\\$,]', '', insurance$BLUEBOOK)
insurance$OLDCLAIM <- gsub('[\\$,]', '', insurance$OLDCLAIM)


insurancetrain <- insurance %>% 
  dplyr::select(-INDEX) %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG),
         KIDSDRIV = as.factor(KIDSDRIV),
         HOMEKIDS = as.factor(HOMEKIDS),
         PARENT1 = as.factor(PARENT1),
         CLM_FREQ = as.factor(CLM_FREQ),
         OLDCLAIM = as.integer(OLDCLAIM),
         BLUEBOOK = as.integer(BLUEBOOK),
         HOME_VAL = as.integer(HOME_VAL),
         INCOME = as.integer(INCOME))

#boxplot, histogram and correlations 
ggplot(melt(insurancetrain), aes(x=factor(variable), y=value)) + facet_wrap(~variable, scale="free") + geom_boxplot()

ggplot(melt(insurancetrain), aes(x=value)) + facet_wrap(~variable, scale="free") + geom_histogram(bins=50)

cor1 <- data.frame(lapply(insurancetrain, function(x) as.numeric(as.factor(x))))

c <- cor(cor1, method="pearson", use="complete.obs")
corrplot(c, method="circle")


```

### We observed that:

* The crime dataset contains 26 variables, with 8161 observations

* There are missing values.


## 2. Data Preparation


```{r}
## checking for no missing data
sapply(insurancetrain, function(x) sum(is.na(x)))

#Using the mice package to input the missing data. 
insurancetraining2 <- mice(insurancetrain, m=5, maxit = 5, method = 'pmm')
insurancetraining2 <- complete(insurancetraining2)
summary(insurancetraining2)
sapply(insurancetraining2, function(x) sum(is.na(x)))


#same for eval set 
sapply(insurance_evaluation, function(x) sum(is.na(x)))
insuranceeval2 <- mice(insurance_evaluation, m=5, maxit = 5, method = 'pmm')
insuranceeval2 <- complete(insuranceeval2)
insuranceeval2 <- data.frame(lapply(insuranceeval2, function(x) as.numeric(as.factor(x))))
summary(insuranceeval2)
sapply(insurancetraining2, function(x) sum(is.na(x)))

```

## 3. Build Models

```{R}
#We will build different multiple linear regression models and binary linear regression models. 
model1 <- lm(TARGET_AMT ~ ., insurancetraining2)
summary(model1)
par(mfrow=c(2,2))
plot(model1)

model2 <- stepAIC(model1, direction = "both", trace = FALSE)
summary(model2)
par(mfrow=c(2,2))
plot(model2)

#box-cox 
insurancebc <- preProcess(insurancetraining2, c("BoxCox"))
insurancebc_transformed <- predict(insurancebc, insurancetraining2)
model4 <- lm(TARGET_AMT ~ ., insurancebc_transformed)
summary(model4)

#For the first three models we see similar results. Where the Q1 and Q3 are not evenly distributed. 
#The r-squared is .29, .28 and .29. Lets look at more models. 

glm_data <- data.frame(lapply(insurancetraining2, function(x) as.numeric(as.factor(x)))) %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) 
glm_data1 <- glm_data %>% 
  dplyr::select(-"TARGET_AMT")

model5 <- glm(TARGET_FLAG ~ ., family = "binomial", glm_data1)
summary(model5)
par(mfrow=c(2,2))
plot(model5)
# Model5 has evenly distributed deviance. many of our variables are significant. 


model6 <- stepAIC(model5, direction = "both", trace = FALSE)
summary(model6)
par(mfrow=c(2,2))
plot(model5)
#This model has similar distribution however the AIC has not improved. 

#box-cox 
glm_data12 <- preProcess(glm_data1, c("BoxCox"))
glmbc_transformed <- predict(glm_data12, glm_data1)
model7 <- glm(TARGET_FLAG ~ ., family = "binomial", glmbc_transformed)
summary(model7)
```



```{R}
# Getting the confusion matix, roc curve for each model 
confusionMatrix1 <- confusionMatrix(as.factor(as.integer(fitted(model5) > .5)), as.factor(model5$y), positive = "1")
rocmodel1 <- roc(glm_data$TARGET_FLAG,  predict(model5, glm_data))
confusionMatrix1
rocmodel1

confusionMatrix2  <- confusionMatrix(as.factor(as.integer(fitted(model6) > .5)), as.factor(model6$y), positive = "1")
rocmodel2 <- roc(glm_data$TARGET_FLAG,  predict(model6, glm_data))
confusionMatrix2
rocmodel2


confusionMatrix3  <- confusionMatrix(as.factor(as.integer(fitted(model7) > .5)), as.factor(model7$y), positive = "1")
rocmodel3 <- roc(glm_data$TARGET_FLAG,  predict(model7, glm_data))
confusionMatrix3
rocmodel3

#Model5 has the highest AUC 
# predict

predict <- predict(model5, insuranceeval2, interval = "prediction")
eval <- table(as.integer(predict > .5))
eval

```
